Finding points on varieties with \emph{Macaulay2}

### 2021-08-19 FastLinAlg/Paper/RandomRationalPoints.tex
Merged the changes of Sankhaneel and Thai

    There is an existing package called {\tt RationalPoints} \cite{RationalPointsSource}, which we took inspiration from, which aims to find \emph{all} the rational points of a variety; our aim here is to find one or more points quickly, even if it the point we find is not rational.  We note also that the package {\tt Cremona} \cite{CremonaSource} can find rational points on projective varieties, as can the core function {\tt randomKRationalPoint} \cite{M2}.  Our methods frequently appear to be faster and apply in the affine setting as well.

@@

%    \vspace{1em}
    \item[\tt NumThreadsToUse => ZZ]

    When calling {\tt randomPoints} with a {\tt BruteForce} strategy, this denotes the number of threads to use in brute force point checking.
    \end{description}
    % Outputs
    % :List
    % a list of points in the variety with possible repetitions.
    % Description
    % Text  
    % Gives at most $n$ many point in a variety $V(I)$. 
    % Example
    % R = ZZ/5[t_1..t_3];
    % I = ideal(t_1,t_2+t_3);
    % randomPoints(3, I)
    % randomPoints(4, I, Strategy => Default)
    % randomPoints(4, I, Strategy => GenericProjection)
    % randomPoints(4, I, Strategy => LinearIntersection)

@@

    %Providing {\tt codimension} speeds up computations significantly as computing codimension using the default command {\tt codim} creates a bottleneck in some cases. If codimension needs to be computed, we  circumvent this issue using a probabilistic code to compute it (see \Cref{helper}). A quick comment, while in many examples generic projection works well, there are some examples where it is many orders of magnitude slower.  By default, generic projection is turned off in the default strategy, that is we set {\tt ProjectionAttempts=>0}.  See \autoref{rem.ProjectingToAHypersurfaceFirst} for additional related discussion.

@@

\subsection{\tt dimViaBezout:} 
\label{subsec.DimViaBezout}
We thank the Frank-Olaf Scheyer for pointing out that in most of the computations, computing the codimension of the given ideal is a significant bottleneck. While we have avoided most dimension computations in our current implementation, we have also implemented a probibalistic dimension computation of $V(I)$.
This function takes as input an ideal $I$ in a polynomial ring over a field and intersects $V(I)$ with random linear spaces of increasing dimension until there is an intersection.  For example, if the intersection of $V(I)$ with a random line has a point, then we expect that $V(I)$ contains a hypersurface.  If there was no intersection, this function tries a $2$-dimensional linear space, and so on.  This can speed up a number of computations.  The function also takes in optional inputs as described below:

@@

\begin{example}
    Consider the following 3-dimensional example (which is regular in codimension 1) where computing the dimension of the singular locus takes around $30$ seconds as there are $15500$ minors of size $4 \times 4$ in the associated $7 \times 12$ Jacobian matrix.  However, we can use this function to quickly find interesting minors.
  {{\small\color{blue}
  \begin{quote}
\begin{verbatim}
		i2 : T = ZZ/101[x1, x2, x3, x4, x5, x6, x7];

>>>>

    There is an existing package called {\tt RationalPoints} \cite{RationalPointsSource}, which we took inspiration from, which aims to find \emph{all} the rational points of a variety; our aim here is to find one or more points quickly, even if it is not rational.  We also note that the package {\tt Cremona} \cite{CremonaSource} can find rational points on projective varieties, as can the core function {\tt randomKRationalPoint} \cite{M2} -- our methods frequently appear to be faster and apply in the affine setting as well.

@@

%    \vspace{1em}
    \item[\tt NumThreadsToUse => ZZ]
When calling  { \tt randomPoints} and functions that call it with a { \tt BruteForce} strategy, this options allows the user to specify the number of threads to use in brute force point checking.

    \end{description}
    % Outputs
    % :List
    % a list of points in the variety with possible repetitions.
    % Description
    % Text  
    % Gives at most $n$ many point in a variety $V(I)$. 
    % Example
    % R = ZZ/5[t_1..t_3];
    % I = ideal(t_1,t_2+t_3);
    % randomPoints(3, I)
    % randomPoints(4, I, Strategy => Default)
    % randomPoints(4, I, Strategy => GenericProjection)
    % randomPoints(4, I, Strategy => LinearIntersection)

@@

    %Providing {\tt codimension} speeds up computations significantly as computing codimension using the default command {\tt codim} creates a bottleneck in some cases. If codimension needs to be computed, we  circumvent this issue using a probabilistic code to compute it (see \Cref{helper}). A quick comment, while in many examples generic projection works well, there are some examples where it is many orders of magnitude slower.  By default, generic projection is turned off in the default strategy, that is we set {\tt ProjectionAttempts=>0}.  See \autoref{rem.ProjectingToAHypersurfaceFirst}  l related discussion.

@@

\subsection{\tt dimViaBezout:} 
\label{subsec.DimViaBezout}
We thank Frank-Olaf Scheyer for pointing out that in most of the computations, computing the codimension of the given ideal is a significant bottleneck. While we have avoided most dimension computations in our current implementation, we have also implemented a probibalistic dimension computation of $V(I)$.
This function takes as input an ideal $I$ in a polynomial ring over a field and intersects $V(I)$ with random linear spaces of increasing dimension until there is an intersection.  For example, if the intersection of $V(I)$ with a random line has a point,  then we expect that $V(I)$ contains a hypersurface.  If there was no intersection, this function tries a $2$-dimensional linear space, and so on.  This can speed up a number of computations.  The function also takes in optional inputs as described below:

@@

\begin{example}
    Consider the following 3-dimensional example which is regular in codimension 1.  Note, in this example, computing the dimension of the singular locus takes around $30$ seconds as there are $15500$ minors of size $4 \times 4$ coming from the associated $7 \times 12$ Jacobian matrix.  However, we can use our function to quickly find interesting minors.
  {{\small\color{blue}
  \begin{quote}
\begin{verbatim}
		i2 : T = ZZ/101[x1, x2, x3, x4, x5, x6, x7];
